\theseminartitle \\ \small \thetitle

### 2021-09-29 tex/paper.tex
Minor changes and corrections to paper

\subsection{Definition}
\label{ssec:bench_definition}
The term "Benchmark" is used in a variety of different domains and, according to the Standard Specialization Evaluation Corporation, historically stems from a physical marking that was used as a reference point on a workbench to check the length of the produced pieces\footnote{SPEC Glossary: \url{https://www.spec.org/spec/glossary/\#benchmark}}. However throughout this paper we are only interested in system benchmarks within the context of software systems and will therefore also refer to the definition of a benchmark as given by Kounev et al.: "a tool coupled with a methodology for the evaluation and comparison of systems or components with respect to specific characteristics, such as performance, reliability or security." (\cite{Kounev} p. 4). As already given from this definition benchmarks can serve as an evaluation for different characteristics. We will more closely look at these in \autoref{ssec:bench_classification}. Throughout this paper we will also mostly refer to the "system or component" that is being benchmarked as the \gls{sut}.

@@

In the broadest sense, benchmarks can be divided into competitive and non-competitive benchmarks. The main motivation of competitive benchmarks is hereby the development of standardized quality criteria that can be used to compare different systems. Certain sources also emphasize this importance on the competitiveness by providing a more detailed definition of the term benchmark, i.e. "a standard tool for competitive evaluation and comparison of competitive systems" (\cite{kistowski2015} p. 333).

For non-competitive benchmarks Kounev et al. Further distinguish between the two categories of "rating tools" and "research benchmark". Rating tools can serve multiple purposes: Either as a component of an development and system improvement process, as a baseline for regulatory programmes or to serve as common benchmark for research. Research benchmarks on the other hand commonly refer to tools and workloads that may be used for in-depth evaluation of research results, prototypes or productive systems. Based on these descriptions it becomes apparent that there sometimes might be no clear distinction between a rating tool and a research benchmark and the terminology is often determined by the evaluating party (See \cite{Kounev} p. 4).

Regardless of the competitiveness, or lack thereof, of a benchmark the key metrics that are evaluated are often similar or even identical. While historically the main metrics of importance were e.g. security, system reliability and energy efficiency, modern benchmarks also try to cover much more aspects of a software. According to Kounev et al. quality metrics of modern benchmarks are usually(\cite{Kounev} p. 6 f):
\begin{itemize}
	\item Performance
	\item Scalability
	\item	Energy Efficiency
	\item Reliability
\end{itemize}

@@

Over the years, researchers and industry associates alike have defined several criteria that are considered desirable by modern standards. Often the exact terminology of these criteria differs based on authors and the specific application domain, however the concepts in general are similar. Kounev at al. list five core criteria that are generally of interest(\cite{Kounev} p. 13f). A much more in depth discussion of these criteria was performed by Kistowski et al. (See \cite{kistowski2015})
\begin{itemize}
	\item Relevance
	\item Reproducibility
	\item Fairness
	\item Verifiability
	\item Usability
\end{itemize}
%\paragraph{Relevance} For a benchmark to be meaningful it is important that it, at least to a certain extent, reproduces some scenarios with real world application. This is what the "Relevance" of a benchmark refers to. It tries to define how relevant the produced information is to a potential user of a software. According to Kistowski et al. this makes it "perhaps the most important characteristic" (\cite{kistowski2015} p. 334). However the relevance of a benchmark may also limits the benchmarks applicability meaning, that a highly relevant benchmark for a specific system or component might only narrow applicability while benchmarks with a broad applicability will have a lower relevance but for a wider selection of applications. Kistowski et al. further elaborate on this with specific examples (\cite{kistowski2015} p. 334f)

@@

	As a result if not including any performance evaluation might lead to performance regressions being shipped in a stable release to customers and users. Therefore using benchmarks throughout the development process can ensure to detect performance changes and allows to fix them before they are being released (See \cite{daly2021} p. 1).

	Additionally employing benchmark on a regular basis also opens the window of opportunity for many other improvements if the corresponding infrastructure is setup accordingly. Daly describes in great detail how this was achieved for the MongoDB infrastructure in \cite{daly2021}. In summary the following have been observed:
	\begin{itemize}
		\item Increased productivity of the developers
		\item Reduced noise on the performance data
		\item Improved Comparison of arbitrary results
		\item An increased amount of performance improvements
	\end{itemize}

@@

	Once the software is nearing a new release and e.g. a new \textit{Release Candidate} is created internally, the development team can use this benchmark to evaluate whether the candidate introduces a performance regression or if it is fit for a release. There are several benefits of this approach; First of all, it does not necessarily requires an automated performance testing infrastructure but rather can be executed and evaluated manually be the development team.

@@

	This can also be observed in commercially used software such as MongoDB which reported about 5500 commits to their code throughout the 2019/2020 release cycle averaging to just about 15 commits per day (\cite{daly2021} p. 2). If the only benchmarking data available is these from the two release candidates this results in an extensive additional overhead to retroactively reconstruct the cause of the root cause. Additionally when only evaluating the performance before release candidates as an affect one would also only detect a performance change at that point. However generally software aims to detect performance defects as soon as possible.

@@

	\subsubsection{Definition}
	As previously discussed in \autoref{ssec:bench_rel} benchmarking only before releases introduces several challenges. The general motivation of continuous benchmarking is to address these by, simply said, generally adding more data points to evaluate. These added data points can then be used for a more detailed analysis of the systems performance over time. The general idea is similar to those used in \gls{ci} and \gls{cd}. By running tests and releases more frequently, often after every commit, these methods aim to shorten release cycles. \gls{cb} however, is a term and methodology that seems not yet commonly defined in academia.

@@

		Using \gls{cb} opens up several opportunities for improvements to the development process of software. First of all it gives a regular overview of a software systems performance. However the usability of these results is also heavily dependent on the quality of the testbed. Especially the use of cloud infrastructure can introduce additional noise to the performance of a system. As has been shown by Daly and Ingo in \cite{daly2019} reducing the noise in such a scenario is a heavily application specific task that can take several weeks of investigating the systems behavior.

	Running performance benchmarks more frequently naturally also enables to detect performance changes more quickly. As a result it also makes it easier to isolate the cause of the change as there may only be fa few commits between the two benchmarks. However again the comparability of these results heavily depends how and if we can take into account the testbed's noise accordingly as to not erroneously flag outliers from noise as actual performance changes. In \cite{daly2021} Daly proposed a technique called \textit{canary testing} to counter these issues. However, in their evaluation canary testing was not viable in their productive system and resulted in more cost than benefit (See \cite{daly2021} p. 6f).

	Another point to consider when using \gls{cb} is the additional data that will inherently get produced. Especially for modern software systems there is not only a single benchmark that is of interest but rather multiple benchmarks with for example different system configurations or workloads. Therefore as Daly et al. and Grambow et al. conclude in their works that manually reviewing the performance data for \gls{cb} is not feasible(See \cite{daly2020} p. 2, \cite{grambow2019} p. 5). As a result, to assist in \gls{cb} and especially in detecting performance changes, automated approaches and toolings are required to pre-process this data and possibly detect performance changes automatically. Approaches for such tools have been described by Daly et al. (\cite{daly2020} p. 2, \cite{daly2021} p. 2ff) and Grambow et al.(\cite{grambow2019}). One of the techniques commonly referred to in this context is change point detection which will be covered in a little more detail in \autoref{sec:cp_detection}.

@@

	\subsection{Definition \& Distinction}
	\label{ssec:cp_distinction}
	The most commonly domain where the term "Change Point" and "Change Point Detection" are commonly used is the one of signal processing. Literature suggest several different definitions of what a change point is.  According to Aminikhanghahi and Cook: "A change point represents a transition between different states in a process that generates the time series data."(\cite{Samaneh2016} p. 4). It is important to understand the fundamental difference between a change point and an outlier in a time series. The NIST\footnote{National Institute of Standards and Technology} Engineering Statistics Handbook defines an outlier as "an observation that lies an abnormal distance from other values in a random sample from a population." (\cite{nist}, part 7.1.6).

@@

	The main idea of the experiment that was created was to perform a benchmark on a series of versions of a specific open-source software to evaluate the methods and techniques described in this paper. More specifically, we want to evaluate the following questions:
	\begin{itemize}
		\item When do we need to benchmark software?
			\begin{itemize}
				\item Is it sufficient to benchmark releases?
				\item Do we need to benchmark every commit?
			\end{itemize}
		\item How do different continuous benchmarking approaches compare with respect to:
			\begin{itemize}
				\item Complexity of implementation?
				\item Reliably detecting performance changes?
				\item Reducing the amount of manual work required by the developer?
			\end{itemize}
	\end{itemize}

	Additionally the experiment should also be used to identify other, currently unidentified, challenges and opportunities that this approach might offer.

	For the first complex of the questions mentioned above, the following methodology is planned: Firstly, perform a benchmark of all public releases on the open-source software that was selected. In a second step then every individual commit is benchmarked for more fine-grained data. By comparing the performance results of the pure release benchmarking and the individual commit benchmarking it is eventually possible to decide if a specific performance regression could have been detected earlier if a continuous benchmarking approach would have been used.

@@

	For the benchmark itself the open source project \texttt{lzbench}(\cite{gitlzbench}) will be used. Lzbench is an in-memory compression benchmarking tool that is designed to compare different implementations of compression algorithms. For this experiment all other compression algorithm implementations will be disabled so that only \texttt{lz4} will be benchmarked at different releases and commits.

	Lzbench offers the option to disable all the other compression algorithms at compile time. The \texttt{lz4} version that is used for the benchmark can simply be changed by overwriting the corresponding lz4 files within the lzbench repository before compilation. In the experiment a slightly adapted version of Lzbench will be used that has the other compression algorithms disabled at compile time and also incorporates some minor changes that were required to ensure compilation on the cluster infrastructure.

	Lzbench also offers a variety of command line switches that influence the behaviour of the benchmark. An overview of the switches that were used for executing this experiment can be found in \autoref{tab:cmd_switches}.

	\begin{table}
		\caption{Overview of used command line switches for lzbench (\cite{gitlzbench})}
		\label{tab:cmd_switches}
		\centering
		\begin{tabular}{|c|c|}
			\hline
			\textbf{Switch} & \textbf{Description}\\
			\hline\hline
			\texttt{-elz4} & Limits the lzbench run to only benchmark the lz4 compression implementation \\
			\hline
			\texttt{-i100,100} & Sets the minimum number of compression and decompression iterations to $100$ \\
			\hline
			\texttt{-p3} & Sets the output to print the median speed of all iterations \\
			\hline
			\texttt{-j} & Join files in memory but compress them independently \\
			\hline
			\texttt{-r} & Operate recursively on directories \\
			\hline 
			\texttt{-o4} & Sets the output text format to CSV \\
			\hline
			\texttt{-v} & Disables progress information \\
			\hline
		\end{tabular}
	\end{table}

@@

	Since the \texttt{lz4} repository consists of over 2500 commits benchmarking each individual commit is infeasible for the scope of this paper as it would both generate too much data to evaluate properly as well as strain the available shared server resources too much. Therefore a two staged approach will be used:
\begin{itemize}
	\item In the first step, a benchmark will be performed on the publicly available releases\footnote{Based on the tags created in the GitHub repository}. The results of these benchmarks will be used to identify a commit range with a significant performance change
	\item All commits in the identified commit range will be benchmarked for a more detailed and thorough analysis.
\end{itemize}

@@

	The benchmark was implemented using a few simple bash scripts that perform an lzbench run on a given list of commits. The benchmark script copies all source files required for building lzbench to a separate build directory. It then checks out the given commit of lz4 and overwrites the lz4 specific files in the lzbench directory. This custom built lzbench version is then used to run the benchmark. All code that was used to perform this benchmark can be found on GitHub (\cite{gitSysBench}).

	\begin{figure}[ht!]
		\centering
		\includegraphics[width=0.8\textwidth]{graph/lz4_releases_plot}
		\caption{Compression- and Decompression speeds throughout the release history}
		\label{fig:release-bench}
	\end{figure}

@@

	Based on these results the commit range between \textit{v1.8.1} and \textit{v1.8.2} was selected as the commit range for a more thorough and detailed analysis. The results of this analysis are further discussed in \autoref{sec:exp_evaluation}. However as the data for the lz4 benchmark only indicated a performance improvement and the goal of this work is to evaluate the detection both performance improvements and regressions, the release benchmarks were repeated with the \texttt{lz4hc} compression algorithm. The \texttt{lz4hc} implementation is a variant of the \texttt{lz4} implementation that features a higher compression rate at the cost of lower compression speed. As \texttt{lz4hc} is also included in the same source code as \texttt{lz4} the only adaption required to the benchmark was to replace the compression algorithm used in the command line parameters passed.

@@

	While the graph for the decompression speed is similar to the one for the normal \texttt{lz4} algorithm, for the compression speed a significant performance regression can be observed between versions \textit{v1.8.0} and \textit{v1.8.1}. Therefore this commit range was also selected for further investigation of the individual commits.

@@

For the analysis of \texttt{lz4} data there are a few key observation that we made while evaluating the data manually. First of all, it seems that there are at least two general "performance levels" that can be observed from the data: While the majority of the data lies around a speed of 2200 MB/s in the end we see a clear improvement to a level of around 2400 MB/s. We also see that at the very beginning and near the end two more steep increases can be oberseved. However as we are missing the data before and after these benchmarks we cannot rule them out to just be outliers in the dataset. Therefore we do not consider these as additional performance levels.

Another observation that we made were the intense performance increases that can be observed throughout the data series that peak to around 2900 MB/s. While manually reviewing the code changes in the corresponding code commits we could not find any connection to the code changes made that would cause such a spike in the decompression speed nor did we find any evidence of changes being reverted in the following commits that would cause the performance to return to the previous level. This observation will be further discussed in \autoref{ssec:refl_quality} and \autoref{ssec:refl_shortcomings}.

@@

\subsection{Comparing Methods to Detect Performance Changes}
After manually reviewing the data we performed a programmatic evaluation of the data with three different methods. Apart from the E-Divisive means algorithm that was already discussed in \autoref{ssec:cp_edivisive}, we will use two more algorithms that were also evaluated by Grambow et al. (See \cite{grambow2019}):

@@

Surprisingly for us, for both datasets the trend detection flagged the most commits as a performance change. We initially expected that because of the many outliers in our dataset that jump detection would flag the most commits as for every outlier two commits would be flagged -- the outlier and the commit right after that. By far the fewest commits were flagged as changes buy the E-Divisive Means algorithm. To our surprise E-Divisive for the \texttt{lz4} algorithm flagged two additional change points with a relatively small overall performance change that we missed in our manual review. While it is impressive that E-Divisive Mean can detect such changes it is a matter of application domain whether detecting such changes is desirable or if further parameters of the algorithm needs to be tuned.

One more aspect we wanted to evaluate was how quickly these techniques can be detect a performance change after it has occured. While our experiment retroactively looks for change points, we are more interested how and if such techniques can be used to find performance changes while they are introduced.

@@

To evaluate how quickly a performance change is detected by E-Divisive Means, we use the data we gather to retroactively "simulate" an ongoing continuous benchmarking approach. That is we iteratively add one commit by one to the time series and therefore find the first occurence of the change points identified in our final time series. The results of this evaluation are shown in \autoref{tab:ediv_occ}

@@

So for our example, E-Divisive Means detects the changes relatively quich in most cases. When reviewing these commits by hand we found that the biggest time difference was about four weeks until the change point got detected (Between commit 104 and 122 for \texttt{lz4hc}). For most of the other change points it was only a few days until the detection. Whether or not a detection time of four weeks is acceptable is heavily project dependent and cannot be answered by us generically.

\subsection{Conclusion}
We believe that all three algorithms presented in this chapter offer useful properties for certain application with respect to detecting performance changes. However when it comes to reliably detecting performance changes throughout the development, we believe that E-Divisive Means stands out against the other two. Jump Detection is neturally prone to noisy signals while for trend detection we observed a fairly poor performance on bi-modal data as we partly oberseved it in our \texttt{lz4} dataset.

However also E-Divisive means has its drawbacks that should be considered before implementing it: For a reliable and vast change detection one needs to regularly benchmark the \gls{sut}, optimally after every code change. In practice this might not be viable. Also we observed that it flags even smaller performance changes which might not desirable. Lastly depending on the performance data it might take some time for the algorithm to identify a change point. However for this last point one needs to consider that E-Divisive Means still identifies the exact point where a performance change was introduced which potentially reduces the amount to find the corresponding code change by far.

@@

\paragraph{Verifiability} -- In terms of verifiability, our benchmark ensures that only a predefined software configuration is used. However our benchmark only copies the build files from an existing directory on the users hardware and does not validate if any unexpected changes have been made. In an extended approach we would rather consider the code to be cloned directly from a public Git repository, preferably with a specific commit defined.

Additionally our detailed experiment contained some outliers that we could not explain based on our findings. We now believe that this might be an error in our benchmarking approach as firstly, the achieved performance results are closely to those from recent \texttt{lz4} versions. Additionally when repeating the benchmark on a smaller subset, we could not reliably reproduce these results. Unfortunately in the time frame of this seminar we could not identify the exact cause of this.

@@

	\subsection{Shortcomings}
	\label{ssec:refl_shortcomings}
	There are several shortcomings of our experiment that can be discussed. First of all we have already discussed that there were some inconsistencies in the data gathered from the experiment. This has caused some, otherwise unexplainable, performance spikes in the data. As we could not reliably reproduce this behavior we believe that there is some issue in our benchmarking pipeline which leads to a wrong version of \texttt{lz4} being used. Therefore, one could argue that it cannot be out ruled that also other data recorded is not flawed in some other unnoticed way.

>>>>

\subsection{Definition}
\label{ssec:bench_definition}
According to the Standard Specialization Evaluation Corporation, historically the term "Benchmark" stems from a physical marking that was used as a reference point on a workbench to check the length of the produced pieces\footnote{SPEC Glossary: \url{https://www.spec.org/spec/glossary/\#benchmark}}. However throughout this paper we are only interested in system benchmarks within the context of software systems and will therefore also refer to the definition of a benchmark as given by Kounev et al.: "a tool coupled with a methodology for the evaluation and comparison of systems or components with respect to specific characteristics, such as performance, reliability or security." (\cite{Kounev} p. 4). As already given from this definition benchmarks can serve as an evaluation for different characteristics. We will more closely look at these in \autoref{ssec:bench_classification}. Throughout this paper we will also mostly refer to the "system or component" that is being benchmarked as the \gls{sut}.

@@

In the broadest sense, benchmarks can be divided into competitive and non-competitive benchmarks. The main motivation of competitive benchmarks is hereby the development of standardized quality criteria that can be used to compare different systems.

For non-competitive benchmarks Kounev et al. Further distinguish between the two categories of "rating tools" and "research benchmark". Rating tools can serve multiple purposes: Either as a component of an development and system improvement process, as a baseline for regulatory programmes or to serve as common benchmark for research. Research benchmarks on the other hand commonly refer to tools and workloads that may be used for in-depth evaluation of research results, prototypes or productive systems. Based on these descriptions it becomes apparent that there sometimes is no clear distinction between a rating tool and a research benchmark and the terminology is often determined by the evaluating party (See \cite{Kounev} p. 4).

Regardless of the competitiveness, or lack thereof, of a benchmark the key metrics that are evaluated are often similar. While historically the main metrics of importance were e.g. security, system reliability and energy efficiency, modern benchmarks also try to cover much more aspects of a software. According to Kounev et al. quality metrics of modern benchmarks are usually(\cite{Kounev} p. 6 f):
\begin{itemize}
	\item Performance
	\item Scalability
	\item	Energy Efficiency
	\item Reliability
\end{itemize}

@@

Over the years, researchers and industry associates alike have defined several criteria that are considered desirable by modern standards. Often the exact terminology of these criteria differs based on authors and the specific application domain, however the concepts in general are similar. Kounev at al. list five core criteria that are generally of interest(\cite{Kounev} p. 13f). A more in depth definition and discussion of these criteria was performed by Kistowski et al. (See \cite{kistowski2015})
\begin{itemize}
	\item Relevance
	\item Reproducibility
	\item Fairness
	\item Verifiability
	\item Usability
\end{itemize}
%\paragraph{Relevance} For a benchmark to be meaningful it is important that it, at least to a certain extent, reproduces some scenarios with real world application. This is what the "Relevance" of a benchmark refers to. It tries to define how relevant the produced information is to a potential user of a software. According to Kistowski et al. this makes it "perhaps the most important characteristic" (\cite{kistowski2015} p. 334). However the relevance of a benchmark may also limits the benchmarks applicability meaning, that a highly relevant benchmark for a specific system or component might only narrow applicability while benchmarks with a broad applicability will have a lower relevance but for a wider selection of applications. Kistowski et al. further elaborate on this with specific examples (\cite{kistowski2015} p. 334f)

@@

	As a result not including any performance evaluation might lead to performance regressions being shipped in a stable release to customers and users. Therefore using benchmarks throughout the development process can ensure to detect performance changes and allows to fix them before they are being released (See \cite{daly2021} p. 1).

	Additionally employing benchmark on a regular basis also opens the window of opportunity for many other improvements if the corresponding infrastructure is setup accordingly. Daly describes in great detail how this was achieved for the MongoDB infrastructure (\cite{daly2021}). In summary the following have been observed:
	\begin{itemize}
		\item Increased productivity of the developers
		\item Reduced noise on the performance data
		\item Improved Comparison of arbitrary results
		\item An increased amount of performance improvements
	\end{itemize}

@@

	Once the software is nearing a new release and e.g. a new \textit{Release Candidate} is created internally, the development team can use this benchmark to evaluate whether the candidate introduces a performance regression or if it is fit for a release. There are several benefits of this approach; First of all, it does not necessarily requires an automated performance testing infrastructure but rather can be executed and evaluated manually by the development team.

@@

	This can also be observed in commercially used software such as MongoDB which reported about 5500 commits to their code throughout the 2019/2020 release cycle averaging to just about 15 commits per day (\cite{daly2021} p. 2). If the only benchmarking data available is these from the two release candidates this results in an extensive additional overhead to retroactively reconstruct the cause of the regression. Additionally when only evaluating the performance before release candidates as an affect one would also only detect a performance change at that point. However generally software aims to detect performance defects as soon as possible.

@@

	\subsubsection{Definition}
	As previously discussed in \autoref{ssec:bench_rel}, benchmarking only before releases introduces several challenges. The general motivation of continuous benchmarking is to address these by, simply said, generating more data points to evaluate. These added data points can then be used for a more detailed analysis of the systems performance over time. The general idea is similar to those used in \gls{ci} and \gls{cd}. By running tests and releases more frequently, often after every commit, these methods aim to shorten release cycles. \gls{cb} however, is a term and methodology that seems not yet commonly defined in academia.

@@

		Using \gls{cb} opens up several opportunities for improvements to the development process of software. First of all it gives a regular overview of a software systems performance. However the usability of these results is also heavily dependent on the quality of the testbed. Especially the use of cloud infrastructure can introduce additional noise to the performance of a system. As has been shown by Daly and Ingo (\cite{daly2019}) reducing the noise in such a scenario is a heavily application specific task that can take several weeks of investigating the systems behavior.

	Running performance benchmarks more frequently naturally also enables to detect performance changes more quickly. As a result it also makes it easier to isolate the cause of the change as there may only be a few commits between the two benchmarks. However again the comparability of these results heavily depends how and if we can take into account the testbed's noise accordingly as to not erroneously flag outliers from noise as actual performance changes. Daly proposed a technique called \textit{canary testing} to counter these issues. However, in their evaluation canary testing was not viable in their productive system and resulted in more cost than benefit (See \cite{daly2021} p. 6f).

	Another point to consider when using \gls{cb} is the additional data that will inherently get produced. Especially for modern software systems there is not only a single benchmark that is of interest but rather multiple benchmarks with for example different system configurations or workloads. Therefore as Daly et al. and Grambow et al. conclude in their works that manually reviewing the performance data for \gls{cb} is not feasible(See \cite{daly2020} p. 2, \cite{grambow2019} p. 5). As a result, to assist in \gls{cb} and especially in detecting performance changes, automated approaches and toolings are required to pre-process this data and possibly detect performance changes automatically. Approaches for such tools have been described by Daly et al. (\cite{daly2020} p. 2, \cite{daly2021} p. 2ff) and Grambow et al.(\cite{grambow2019}). One of the techniques commonly referred to in this context is change point detection which will be covered in more detail in \autoref{sec:cp_detection}.

@@

	\subsection{Definition \& Distinction}
	\label{ssec:cp_distinction}
	The domain where the terms "Change Point" and "Change Point Detection" are commonly used is the one of signal processing. Literature suggest several different definitions of what a change point is.  According to Aminikhanghahi and Cook: "A change point represents a transition between different states in a process that generates the time series data."(\cite{Samaneh2016} p. 4). It is important to understand the fundamental difference between a change point and an outlier in a time series. The NIST\footnote{National Institute of Standards and Technology} Engineering Statistics Handbook defines an outlier as "an observation that lies an abnormal distance from other values in a random sample from a population." (\cite{nist}, part 7.1.6).

@@

	The main idea of the experiment performed was to perform a benchmark on a series of versions of a specific open-source software to evaluate the methods and techniques described in this paper. More specifically, we want to evaluate the following questions:
	\begin{itemize}
		\item When do we need to benchmark software?
			\begin{itemize}
				\item Is it sufficient to benchmark releases?
				\item Do we need to benchmark every commit?
			\end{itemize}
		\item How do different continuous benchmarking approaches compare with respect to:
			\begin{itemize}
				\item Complexity of implementation?
				\item Reliably detecting performance changes?
				\item Reducing the amount of manual work required by the developer?
			\end{itemize}
	\end{itemize}

	Additionally the experiment should also be used to identify other, currently unidentified, challenges and opportunities that this approach might pose.

	For the first complex of the questions mentioned above, the following methodology is planned: First, perform a benchmark of all public releases on the open-source software that was selected. In a second step then every individual commit is benchmarked for more fine-grained data. By comparing the performance results of the pure release benchmarking and the individual commit benchmarking it is eventually possible to decide if a specific performance regression could have been detected earlier if a continuous benchmarking approach would have been used.

@@

	For the benchmark itself the open source project \texttt{lzbench}(\cite{gitlzbench}) will be used. \texttt{Lzbench} is an in-memory compression benchmarking tool that is designed to compare different implementations of compression algorithms. For this experiment all other compression algorithm implementations will be disabled so that only \texttt{lz4} will be benchmarked at different releases and commits.

	\texttt{Lzbench} offers the option to disable all the other compression algorithms at compile time. The \texttt{lz4} version that is used for the benchmark can simply be changed by overwriting the corresponding \texttt{lz4} files within the \texttt{lzbench} repository before compilation. In the experiment a slightly adapted version of Lzbench will be used that has the other compression algorithms disabled at compile time and also incorporates some minor changes that were required to ensure compilation on the cluster infrastructure.

	\texttt{Lzbench} also offers a variety of command line switches that influence the behaviour of the benchmark. An overview of the switches that were used for executing this experiment can be found in \autoref{tab:cmd_switches}.

	\begin{table}
		\caption{Overview of used command line switches for lzbench (\cite{gitlzbench})}
		\label{tab:cmd_switches}
		\centering
		\begin{tabular}{|c|c|}
			\hline
			\textbf{Switch} & \textbf{Description}\\
			\hline\hline
			\texttt{-elz4} & Limits the lzbench run to only benchmark the lz4 compression implementation \\
			\hline
			\texttt{-t60,60} & Sets the minimum time for compression and decompression to $60$ seconds \\
			\hline
			\texttt{-p3} & Sets the output to print the median speed of all iterations \\
			\hline
			\texttt{-j} & Join files in memory but compress them independently \\
			\hline
			\texttt{-r} & Operate recursively on directories \\
			\hline 
			\texttt{-o4} & Sets the output text format to CSV \\
			\hline
			\texttt{-v} & Disables progress information \\
			\hline
		\end{tabular}
	\end{table}

@@

	Since the \texttt{lz4} repository consists of over 2500 commits benchmarking each individual commit is infeasible for the scope of this paper as it would both generate too much data to evaluate properly as well as strain the available shared server resources too much. Therefore a two staged approach will be used:
\begin{itemize}
	\item In the first step, a benchmark will be performed on the publicly available releases\footnote{Based on the tags created in the GitHub repository}. The results of these benchmarks will be used to identify a commit range with a significant performance change
	\item All commits in the identified commit range will be benchmarked for a more detailed analysis.
\end{itemize}

@@

	The benchmark was implemented using a few simple bash scripts that performs an \texttt{lzbench} run on a given list of commits. The benchmark script copies all source files required for building \texttt{lzbench} to a separate build directory. It then checks out the given commit of lz4 and overwrites the \texttt{lz4} specific files in the \texttt{lzbench} directory. This custom built \texttt{lzbench} version is then used to run the benchmark. All code that was used to perform this benchmark can be found on GitHub (\cite{gitSysBench}).

	\begin{figure}[ht!]
		\centering
		\includegraphics[width=0.8\textwidth]{graph/lz4_releases_plot}
		\caption{Compression- and Decompression speeds of throughout the release history}
		\label{fig:release-bench}
	\end{figure}

@@

@@

	Based on these results the commit range between \textit{v1.8.1} and \textit{v1.8.2} was selected as the commit range for a more thorough and detailed analysis. The results of this analysis are further discussed in \autoref{sec:exp_evaluation}. However as the data for the \texttt{lz4} benchmark only indicated a performance improvement and the goal of this work is to evaluate the detection both performance improvements and regressions, the release benchmarks were repeated with the \texttt{lz4hc} compression algorithm. The \texttt{lz4hc} implementation is a variant of the \texttt{lz4} implementation that features a higher compression rate at the cost of lower compression speed. As \texttt{lz4hc} is also included in the same source code as \texttt{lz4} the only adaption required to the benchmark was to replace the compression algorithm used in the command line parameters passed.

	The results of this analysis are shown in \autoref{fig:release-bench-lz4hc}. While the graph for the decompression speed is similar to the one for the normal \texttt{lz4} algorithm, for the compression speed a significant performance regression can be observed between versions \textit{v1.8.0} and \textit{v1.8.1}. Therefore this commit range was also selected for further investigation of the individual commits.

@@

For the analysis of \texttt{lz4} data there are a few key observation that we made while evaluating the data manually. First of all, it seems that there are at least two general "performance levels" that can be observed from the data: While the majority of the data lies around a speed of 2200 MB/s in the end we see a clear improvement to a level of around 2400 MB/s. We also see that at the very beginning and near the end two more steep increases can be observed. However as we are missing the data before and after these benchmarks we cannot rule them out to just be outliers in the dataset. Therefore we do not consider these as additional performance levels.

Another observation that we made were the abrupt performance increases that can be observed throughout the data series that peak to around 2900 MB/s. While manually reviewing the code changes in the corresponding code commits we could not find any connection to the code changes made that would cause such a spike in the decompression speed nor did we find any evidence of changes being reverted in the following commits that would cause the performance to return to the previous level. This observation will be further discussed in \autoref{ssec:refl_quality} and \autoref{ssec:refl_shortcomings}.

@@

\subsection{Comparing Methods to Detect Performance Changes}
After manually reviewing the data we performed a programmatic evaluation of the data with three different methods. Apart from the E-Divisive means algorithm that was already discussed in \autoref{ssec:cp_edivisive}, we will use two more algorithms that were also evaluated by Grambow et al. (See \cite{grambow2019} p. 3):

@@

Surprisingly for us, for both datasets the trend detection flagged the most commits as a performance change. We initially expected, because of the many outliers in our dataset, that jump detection would flag the most commits as for every outlier two commits would be flagged -- the outlier and the commit right after that. By far the fewest commits were flagged as changes buy the E-Divisive Means algorithm. To our surprise E-Divisive for the \texttt{lz4} algorithm flagged two additional change points with a relatively small overall performance change that we missed in our manual review. While it is interesting that E-Divisive Mean can detect such changes it is a matter of application domain whether detecting such changes is desirable or if further parameters of the algorithm needs to be tuned.

One more aspect we wanted to evaluate was how quickly these techniques can detect a performance change after it has occurred. While our experiment retroactively looks for change points, we are more interested how and if such techniques can be used to find performance changes while they are introduced.

@@

To evaluate how quickly a performance change is detected by E-Divisive Means, we use the data we gathered to retroactively "simulate" an ongoing continuous benchmarking approach. That is we iteratively add one commit by one to the time series and find the first occurrence of the change points identified in our final time series. The results of this evaluation are shown in \autoref{tab:ediv_occ}

@@

For our example, E-Divisive Means detects the changes relatively quick in most cases. When reviewing these commits by hand we found that the biggest time difference was about four weeks until the change point got detected (Between commit 104 and 122 for \texttt{lz4hc}). For most of the other change points it was only a few days until the detection. Whether or not a detection time of four weeks is acceptable is heavily project dependent and cannot be answered by us generically.

\subsection{Conclusion}
We believe that all three algorithms presented in this chapter offer useful properties for certain application with respect to detecting performance changes. However when it comes to reliably detecting performance changes throughout the development, we believe that E-Divisive Means stands out against the other two. Jump Detection is naturally prone to noisy signals while for trend detection we observed a fairly poor performance on bi-modal data as we partly observed it in our \texttt{lz4} dataset.

However also E-Divisive means has its drawbacks that should be considered before implementing it: For a reliable and vast change detection one needs to regularly benchmark the \gls{sut}, optimally after every code change. In practice this might not be viable. Also we observed that it flags even smaller performance changes which might not desirable. Lastly depending on the performance data it might take some time for the algorithm to identify a change point. However for this last point one needs to consider that E-Divisive Means still identifies the exact point where a performance change was introduced which potentially reduces the amount to find the corresponding code change.

@@

\paragraph{Verifiability} -- In terms of verifiability, our benchmark ensures that only a predefined software configuration is used. However our benchmark only copies the build files from an existing directory on the users hardware and does not validate if any unexpected changes have been made. In an extended approach we would rather consider the code to be cloned directly from the public Git repository, preferably with a specific commit defined.

Additionally our detailed experiment contained some outliers that we could not explain based on our findings. We now believe that this might be an error in our benchmarking approach as the achieved performance results are closely to those from recent \texttt{lz4} versions. Additionally when repeating the benchmark on a smaller subset, we could not reliably reproduce these results. Unfortunately in the time frame of this seminar we could not identify the exact cause of this.

@@

	\subsection{Shortcomings}
	\label{ssec:refl_shortcomings}
	There are several shortcomings of our experiment that can be discussed. First of all we have already discussed that there were some inconsistencies in the data gathered from the experiment. This has caused some, otherwise unexplainable, performance spikes in the data. As we could not reliably reproduce this behavior we believe that there is some issue in our benchmarking pipeline which leads to a wrong version of \texttt{lz4} being used. Therefore, one could argue that it cannot be ruled out that also other data recorded is not flawed in some other unnoticed way.
